---
title: "p8105_hw2_xz3298"
author: "xinyi zheng"
date: "2023-10-04"
output: github_document
---

##xz3298
##Xinyi Zheng

##problem 2
```{r}
library(readxl)
library(dplyr)
library(janitor)
```

# import and clean Mr. Trash Wheel's data.
```{r}
mrtrash_df=
  readxl::read_excel("/Users/zhengxinyi/Xinyi77/data/202309 Trash Wheel Collection Data.xlsx", sheet= "Mr. Trash Wheel", range = "A2:N550", skip = 2)|>
  janitor::clean_names()|>
  filter(!is.na(dumpster))|>
  mutate(homes_powered=weight_tons * 500/30,
         trash_wheel="mr",
         year=as.numeric(year))
```

# rename columns for clarity
```{r}
colnames(mrtrash_df) <- c("dumpster", "month", "year", "date", "weight", "volume", "plastic bottles", "polystyrene", "cigarette", "glass bottles", "grocery bags", "chip bags", "sports balls", "home powered")
```

# clean Professor Trash Wheel.
```{r}
proftrash_df=
  readxl::read_excel("/Users/zhengxinyi/Xinyi77/data/202309 Trash Wheel Collection Data.xlsx", sheet= "Professor Trash Wheel", range = "A2:M97", skip = 2)|>
  janitor::clean_names()|>
  filter(!is.na(dumpster))|>
  mutate(homes_powered=weight_tons * 500/30,
         trash_wheel="professor")
```

# rename columns for clarity
```{r}
colnames(proftrash_df) <- c("dumpster", "month", "year", "date", "weight", "volume", "plastic bottles", "polystyrene", "cigarette", "glass bottles", "grocery bags", "chip bags", "sports balls", "home powered")
```

# clean Gwynnda Trash Wheel
```{r}
gwytrash_df=
  readxl::read_excel("/Users/zhengxinyi/Xinyi77/data/202309 Trash Wheel Collection Data.xlsx", sheet= "Gwynnda Trash Wheel", range = "A2:K110", skip = 2)|>
  janitor::clean_names()|>
  filter(!is.na(dumpster))|>
  mutate(homes_powered=weight_tons * 500/30,
         trash_wheel="gwynnda")
```

# rename columns for clarity
```{r}
colnames(gwytrash_df) <- c("dumpster", "month", "year", "date", "weight", "volume", "plastic bottles", "polystyrene", "cigarette", "glass bottles", "grocery bags", "chip bags", "sports balls", "home powered")
```

# combine datasets.
```{r}
mrtrash_df$year <- as.character(mrtrash_df$year)
proftrash_df$year <- as.character(proftrash_df$year)
gwytrash_df$year <- as.character(gwytrash_df$year)

all_data <- bind_rows(mrtrash_df, proftrash_df, gwytrash_df)
```

# Dataset Descriptions:

1. mrtrash_df:
    Number of Observations: This dataset contains `r nrow(mrtrash_df)` observations.
    Key Variables: The main variables of interest in this dataset are `dumpster`, `weight_tons`, and `homes_powered`.

2. proftrash_df:
    Number of Observations: There are `r nrow(proftrash_df)` observations in this dataset.
    Key Variables: This dataset primarily focuses on variables such as `dumpster`, `weight_tons`, and `homes_powered`.

3. gwytrash_df:
    Number of Observations: This dataset comprises `r nrow(gwytrash_df)` observations.
    Key Variables: Notably, the dataset includes `cigarette_butts`, `weight`, and `homes_powered` as its main variables.

4. all_data:
    Structure: The combined dataset, `all_data`, consists of `r ncol(all_data)` variables and `r nrow(all_data)` observations.
    Key Variables: The dataset encompasses various attributes, with primary ones being the types of trash wheel (`Mr`, `Professor`, or `Gwynnda`), `weight_tons`, `cigarette_butts`, and other similar items.

In essence, these descriptions give a brief overview of the data's structure and highlight the major variables in each dataset. The `r` notation is indicative of inline R code, which would be executed when rendering the document in an R Markdown environment to display the actual numeric values.
```

# Total weight of trash collected by Professor Trash Wheel
```{r}
prof_total_weight <- sum(proftrash_df$weight, na.rm = TRUE)
print(prof_total_weight)
```
# Total number of cigarette butts collected by Gwynnda in July of 2021
```{r}
gwynnda_cigarettes_july_2021 <- sum(gwytrash_df[gwytrash_df$month == "July" & gwytrash_df$year == "2021",]$cigarette, na.rm = TRUE)
print(gwynnda_cigarettes_july_2021)
```


##problem 3
```{r}
library(readxl)
library(dplyr)
library(janitor)
```
# import and clean the baseline data
```{r, warning=FALSE}
baseline = read.csv("/Users/zhengxinyi/Xinyi77/data/MCI_baseline.csv", skip = 1) %>%
  janitor::clean_names() %>%
  mutate(sex = ifelse(sex == 1, "male", "female"),
         apoe4 = ifelse(apoe4 == 1, "carrier", "non_carrier"),
         age_at_onset = as.numeric(age_at_onset))

baseline_mci = baseline %>%
  drop_na(age_at_onset)

baseline <- baseline %>% filter(MCI_status != "Yes")
```
# import and clean the biomarker data
```{r, warning=FALSE}
biomarker = read.csv("/Users/zhengxinyi/Xinyi77/data/mci_amyloid.csv", skip = 1) %>%
  janitor::clean_names() %>%
  rename(id = study_id,
         time_0 = baseline) %>%
  pivot_longer(
    time_0:time_8, 
    names_to = "time_in_years",
    values_to = "amyloid_ratio"
  )  %>%
  mutate(amyloid_ratio = as.numeric(amyloid_ratio))
```
# analysis
```{r}
total_participants <- nrow(baseline)
mci_participants <- sum(!is.na(baseline$Age_at_onset))
avg_age <- mean(baseline$Current_Age, na.rm = TRUE)
female_apoe4 <- mean(baseline$Sex == 0 & baseline$apoe4 == 1, na.rm = TRUE)
```
# merge data
```{r}
combined_data <- inner_join(baseline, biomarker, by = c("ID" = "Study.ID"))
```
# check the missing
```{r}
missing_ids <- combined_data$ID[is.na(combined_data$Current_Age) | is.na(combined_data$Amyloid_42_40_Ratio)]
```
# export data
```{r}
write.csv(combined_data, "combined_data.csv", row.names = FALSE)
```

## Descriptions:

1. baseline:
    Number of Observations: This dataset, sourced from 'MCI_baseline.csv', contains `r nrow(baseline)` observations.
    Key Variables: Notable variables in this dataset include `sex`, `apoe4`, and `age_at_onset`. The `sex` variable is categorized as either "male" or "female", while the `apoe4` variable is classified as "carrier" or "non_carrier".
    Additional Note: The data has been filtered to exclude observations where the `MCI_status` is "Yes".

2. biomarker:
    Number of Observations: Sourced from 'mci_amyloid.csv', this dataset comprises `r nrow(biomarker)` observations.
    Key Variables: Central variables in this dataset are `time_in_years` and `amyloid_ratio`. The data has been reshaped to have a long format for the time and amyloid ratio variables.

3. combined_data:
    Number of Observations: The dataset formed by merging `baseline` and `biomarker` datasets has `r nrow(combined_data)` observations.
    Key Variables: The `combined_data` dataset contains key attributes from both the `baseline` and `biomarker` datasets, including variables such as `sex`, `apoe4`, `time_in_years`, and `amyloid_ratio`.
    Additional Note: This merged dataset has been saved as 'combined_data.csv'.

4. General Analysis:
    The total number of participants in the `baseline` dataset is `r total_participants`.
    The number of participants with a non-missing 'Age_at_onset' value in the `baseline` dataset is `r mci_participants`.
    The average current age of participants from the `baseline` dataset is `r avg_age`.
    The proportion of participants in the `baseline` dataset who are female and are apoe4 carriers is `r female_apoe4`.

5. Missing Data:
    The dataset contains certain IDs with missing values either for the `Current_Age` or the `Amyloid_42_40_Ratio`, and these IDs are stored in the `missing_ids` variable.
